---
title: "Assignment 5"
author: "Sungjoo Cho"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub link:
https://github.com/katelmrx/cho-lamoreaux-a1.git

## Packages
```{r message=FALSE, warning=FALSE}
library(censusapi)
library(tidyverse)
library(magrittr)
library(factoextra)
library(base)
library(RSocrata)
library(ggmap)
library(lubridate)
library(corrplot)
library(dplyr)
library(knitr)
```


## Exploring ACS Data

```{r include=FALSE}
cs_key <- "839176755fb51cd99853fa546eac00eb08b203fa"
```


```{r}
# Data from the American Community Survey (ACS)
acs_il_c <- getCensus(name = "acs/acs5",
                      vintage = 2016,
                      vars = c("NAME", "B01003_001E", "B19013_001E", "B19301_001E"),
                      region = "county:*",
                      regionin = "state:17",
                      key = cs_key) %>%
  rename(pop = B01003_001E,
         hh_income = B19013_001E,
         income = B19301_001E)
head(acs_il_c)
```

```{r}
# Pull map data for Illinois into a data frame.
il_map <- map_data("county", region = "illinois")
head(il_map)
```

Join the ACS data with the map data. Not that il_map has a column subregion which includes county names. We need a corresponding variable in the ACS data to join both data sets. This needs some transformations, among which the function tolower() might be useful. Call the joined data acs_map.

```{r}
# creating county_name variable in acs_il
acs_il_c$subregion <- gsub(", .*", "", acs_il_c$NAME)
acs_il_c$subregion <- gsub("County", "", acs_il_c$subregion)
acs_il_c$subregion <- trimws(acs_il_c$subregion, "right")
acs_il_c$subregion <- tolower(acs_il_c$subregion)

# join
acs_map <- inner_join(acs_il_c, il_map, by=join_by("subregion" == "subregion"))
head(acs_map)

# number of rows
dim(acs_il_c)
dim(il_map)
dim(acs_map)
```

After you do this, plot a map of Illinois with Counties colored by per capita income.
```{r}
#| fig.height = 6.5, fig.width = 6

ggplot(acs_map) + 
geom_polygon(aes(x = long, y = lat, group = group, fill = income))
```

## Hierarchical Clustering

We want to find clusters of counties that are similar in their population, average household income and per capita income. First, clean the data so that you have the appropriate variables to use for clustering. 
```{r}
head(acs_map)
# select pop, hh_income, and income variables
hclust_data <- acs_map %>%
  select(pop, hh_income, income)

# cleaned data
head(hclust_data)
```

Next, create the distance matrix of the cleaned data. This distance matrix can be used to cluster counties, e.g. using the ward method.
```{r}
# distance matrix of the cleaned data
hclust_d <- dist(hclust_data)
as.matrix(hclust_data)

# clustering counties using the ward method
hc_ward <- hclust(hclust_d, method = "ward.D2")

# clustering counties using the complete method
hc_complete <- hclust(hclust_d, method = "complete")

# clustering counties using the average method
hc_average <- hclust(hclust_d, method = "average")
```


Plot the dendrogram to find a reasonable number of clusters. Draw boxes around the clusters of your cluster solution.
```{r}
# ward method
plot(hc_ward, main = "Ward", xlab = "", sub = "")

# complete method
plot(hc_complete, main = "Complete Linkage", xlab = "", sub = "")

# average method
plot(hc_average, main = "Average Linkage", xlab = "", sub = "")

rect.hclust(hc_ward, 
            k = 5, 
            border = "red")
```

We want to create four clusters based on the `hc_ward` object.
```{r}
cutree(hc_ward, 5)
```

Visualize the county clusters on a map. For this task, create a new `acs_map` object that now also includes cluster membership as a new column. This column should be called `cluster`.
```{r}
# adding cluster membership variable
acs_map <- acs_map %>%
  mutate(cluster = cutree(hc_ward, 5))

# new acs_map
head(acs_map)

# visualize the county clusters on map
ggplot(acs_map) +geom_polygon(aes(x=long, y=lat, group=group, fill=factor(cluster))) +
  theme_minimal( ) + labs(fill="cluster")
```


## Census Tracts

For the next section we need ACS data on a census tract level. We use the same variables as before.

```{r}
acs_il_t <- getCensus(name = "acs/acs5",
                      vintage = 2016,
                      vars = c("NAME", "B01003_001E", "B19013_001E", "B19301_001E"),
                      region = "tract:*",
                      regionin = "state:17",
                      key = cs_key) %>%
  mutate_all(list(~ ifelse(.==-666666666, NA, .))) %>%
  rename(pop = B01003_001E,
         hh_income = B19013_001E,
         income = B19301_001E)

acs_il_t <- na.omit(acs_il_t)
head(acs_il_t)
```

## k-Means

As before, clean our data for clustering census tracts based on population, average household income and per capita income.

```{r}
# select pop, hh_income, income
acs_il_t_cleaned <- acs_il_t %>%
  select(pop, hh_income, income) %>%
  na.omit()

# new dataset
head(acs_il_t_cleaned)
```

Since we want to use K Means in this section, we start by determining the optimal number of K that results in Clusters with low within but high between variation. Plot within cluster sums of squares for a range of K (e.g. up to 20).

* We choose k = 6 as the optimal number of clusters.

```{r}
# finding the optimal number of K
fviz_nbclust(acs_il_t_cleaned, kmeans, method = "wss", k.max = 20)
```

Run `kmeans()` for the optimal number of clusters based on the plot above.
```{r}
# kmeans
km <- kmeans(acs_il_t_cleaned, 6, nstart = 20)
km
```


Find the mean population, household income and per capita income grouped by clusters. In addition, display the most frequent county that can be observed within each cluster.
```{r}
# adding cluster variable
acs_il_t_cluster <- acs_il_t %>%
  mutate(cluster = km$cluster)

# mean population, household income and per capita income grouped by clusters
stat <- acs_il_t_cluster %>%
  group_by(cluster) %>%
  summarize(mean_pop = mean(pop),
            mean_hh_income = mean(hh_income),
            mean_income = mean(income),
            most_frequent_county = names(which.max(table(NAME))))

# table
kable(stat, caption = "Statistics")
```

As you might have seen earlier, it's not always clear which number of clusters is the optimal choice. To automate K Means clustering, program a function based on `kmeans()` that takes K as an argument. You can fix the other arguments, e.g. such that a specific dataset is always used when calling the function.


We want to utilize this function to iterate over multiple Ks (e.g., K = 2, ..., 10) and \-- each time \-- add the resulting cluster membership as a new variable to our (cleaned) original data frame (`acs_il_t`). There are multiple solutions for this task, e.g. think about the `apply` family or `for` loops.
```{r}
# create function
function_kmeans <- function(data, K) {
  cluster_data <- data %>%
    select(pop, hh_income, income)
  kmeans_result <- kmeans(cluster_data, K)
  acs_il_t <- acs_il_t %>%
    mutate(K = as.factor(kmeans_result$cluster))
  return(acs_il_t)
}

# K values
Ks <- 2:10

for (K in Ks) {
  function_kmeans(acs_il_t_cleaned, K)
}
```

Finally, display the first rows of the updated data set (with multiple cluster columns).
```{r}
head(acs_il_t)
```






